Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-01-07 manuscript.Rnw
More tweaking to citations and bibliography.
more citations

\bibliographystyle{unsrtnat}
\linenumbers

@@

Globally, forests store ca. 360 Pg C in live biomass \citep{panetal2011}, or about 40 years of current athropogenic emissions \citep{friedlingsteinetal2014}. The partitioning of this biomass among plant components is a major determinant of vegetation productivity and the rate of carbon cycling. Biomass stored in woody stems has a long residence time \citep{luyssaert2008},  thus representing a significant long-term carbon storage component. By contrast, leaf biomass turns over quickly, and is a major input to the soil carbon cycle where the majority of carbon is released back to the atmosphere. Recently, the residence time of carbon in vegetation has been identified as a major source of uncertainty in model projections of effects of climate change on the global carbon sink \citep{friend2014}. There is little consensus in these global vegetation models (GVMs) on how biomass partitioning should be modelled \citep{franklin2012,dekauwe2014}, but there is no shortage of underlying theory \citep{cannelldewar1994,franklin2012}. We stress that biomass partitioning, the actual distribution of biomass among components, is distinct from biomass allocation, the flux of carbon into plant components. Biomass partitioning is estimated from allocation by including the rates of turnover of the various plant components (see Methods S1).

@@

We found that PFTs differed in LMF, and this was proportional to leaf mass per area (LMA), a leading trait within the leaf economics spectrum. In addition, LMA alone explained 28\% of the variation in $M_F/A_S$ across species, without accounting for PFT (Fig. S6) (and consistent with averages by PFT, $A_F/A_S$ was very weakly related to LMA, with an $R^2$ of 0.02). It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. One explanation for this proportionality may be the correlation between LMA and leaf lifespan \citep{wrightetal2004}. If annual leaf production is constant, longer leaf lifespan will simply result in the accumulation of more leaf cohorts, and thus biomass \citep{enquistniklas2002}. However, leaf lifespan (LL) is correlated to but not proportional to LMA ($LL \propto LMA^{0.58}$, from Ref. \citep{wrightetal2004}). Another interpretation is that plants may target a certain leaf area given a certain biomass investment in supporting tissues, because light interception and photosynthesis correlate with the exchange area of the plant.

>>>>

\bibliographystyle{ieeetr}
\linenumbers

@@

Globally, forests store ca. 360 Pg C in live biomass \citep{panetal2011}, or about 40 years of current athropogenic emissions \citep{friedlingsteinetal2014}. The partitioning of this biomass among plant components is a major determinant of vegetation productivity and the rate of carbon cycling \cite{ise2010}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008},  thus representing a significant long-term carbon storage component. By contrast, leaf biomass turns over quickly, and is a major input to the soil carbon cycle where the majority of carbon is released back to the atmosphere. Recently, the residence time of carbon in vegetation has been identified as a major source of uncertainty in model projections of effects of climate change on the global carbon sink \citep{friend2014}. There is little consensus in these global vegetation models (GVMs) on how biomass partitioning should be modelled \citep{franklin2012,dekauwe2014}, but there is no shortage of underlying theory \citep{cannelldewar1994,franklin2012}. We stress that biomass partitioning, the actual distribution of biomass among components, is distinct from biomass allocation, the flux of carbon into plant components. Biomass partitioning is estimated from allocation by including the rates of turnover of the various plant components (see Methods S1).

@@

We found that PFTs differed in LMF, and this was proportional to leaf mass per area (LMA), a leading trait within the leaf economics spectrum. In addition, LMA alone explained 28\% of the variation in $M_F/A_S$ across species, without accounting for PFT (Fig. S6) (and consistent with averages by PFT, $A_F/A_S$ was very weakly related to LMA, with an $R^2$ of 0.02). It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. One explanation for this proportionality may be the correlation between LMA and leaf lifespan \citep{wrightetal2004}. If annual leaf production is constant, longer leaf lifespan will simply result in the accumulation of more leaf cohorts, and thus biomass \citep{chabothicks1982,enquistniklas2002,poorter2012}. However, leaf lifespan (LL) is correlated to but not proportional to LMA ($LL \propto LMA^{0.58}$, from Ref. \citep{wrightetal2004}). Another interpretation is that plants may target a certain leaf area given a certain biomass investment in supporting tissues, because light interception and photosynthesis correlate with the exchange area of the plant.
